Supporting Information

### 2015-08-19 manuscript_supportinginfo.Rmd
Many minor fixes

**Figure S1**. Increase in gapfraction ($\tau_d$) and total litter production during the 4-month drought period in 2013, calculated for each ring separately. Blue colours are ambient, red colours elevated C~a~. (a) The change in  $\tau_d$ and its 95% confidence interval was calculated from the slope of linear regression of $\tau_d$ with time (number of days since 8 July 2013). (b) Mean litter production expressed in terms of LAI for each ring. The confidence interval was calculated based on the 8 litter traps in each ring. Note the overlapping confidence intervals for all rings.

@@

We took advantage of the detailed timeseries of LAI estimates from canopy transmittance to test a simple photography-based method proposed by Macfarlane et al. [-@Macfarlane2007], which is extensively applied to *Eucalyptus*-dominated forests. From October 2012 to October 2013, we took ca. monthly canopy photographs in each FACE ring. Photographs were taken in the morning on cloudy days. We took one photograph at each point of a 4 x 4 m grid established within each ring, unless obstructed by trees, yielding 20-23 photos per ring. Sampling points too close to either the vertical vent pipes (edges of the ring) or the central scaffolding tower were excluded. The camera was aligned with the north-south axis and set at 1.25 m height.  The camera lens was pointed directly upwards and leveled using a three axis bubble level. We used a Nikon D90 camera with a Nikon AF 50 mm lens (Nikon Corporation, Tokyo, Japan). To measure cover, the camera was set to automatic exposure, Aperture-Priority and automatic autofocus mode and F10 lens. Following recommendations by Jonckheere et al. [-@Jonckheere2005], we used the automated Ridler algorithm to threshold the canopy photos. We then calculated canopy transmittance ($\tau_p$) from each canopy photo as the proportion of pixels with intensity in the blue band below that threshold. LAI from canopy photos ($L_p$) was calculated with a simple Beer-Lambert law (Eq. 4).

@@

  where $\tau_p$ canopy transmission estimated from canopy photos. We assumed that the extinction coefficient *k* = 0.5, following Macfarlane et al. [-@Macfarlane2007] recommendation for eucalypt forests. Without further calibration, the two estimates were reasonably well correlated, though on average $L_\tau$ was `r f(100*(with(flatcan_byring, mean(LAI.PAR.mean / LAI))-1))`% higher than $L_p$ (R^2^ = `r f(glance(lmphoto)$r.squared)`, RMSE = `r f(glance(lmphoto)$sigma)`) (Fig. S2). We also estimated the extinction coefficient that minimized the difference between the two estimates of LAI (Fig. S2), giving *k* = `r f(kopt)`. These comparisons are encouraging because the photos measure LAI near the zenith (+/- 15^o^), while $L_\tau$ is integrated over the sky hemisphere (Eq. 2), weighted towards the zenith.

>>>>

**Figure S1**. Increase in gapfraction ($\tau_d$) and total litter production during the 4-month drought period in 2013, calculated for each ring separately. Blue colours are ambient (nrs 2,3,6), red colours elevated C~a~ (nrs 1,4,5). (a) The change in  $\tau_d$ and its 95% confidence interval was calculated from the slope of linear regression of $\tau_d$ with time (number of days since 8 July 2013). (b) Total litter production during the drought expressed in terms of LAI for each ring. The confidence interval was calculated based on the 8 litter traps in each ring. Note the overlapping confidence intervals for all rings.

@@

We took advantage of the detailed timeseries of LAI estimates from canopy transmittance to test a simple photography-based method proposed by Macfarlane et al. [-@Macfarlane2007], which is extensively applied to *Eucalyptus*-dominated forests. From October 2012 to October 2013, we took ca. monthly canopy photographs in each FACE ring. Photographs were taken in the morning on cloudy days. We took one photograph at each point of a 4 x 4 m grid established within each ring, unless obstructed by trees, yielding 20-23 photos per ring. Sampling points too close to either the vertical vent pipes (edges of the ring) or the central scaffolding tower were excluded. The camera was aligned with the north-south axis and set at 1.25 m height.  The camera lens was pointed directly upwards and leveled using a three axis bubble level. We used a Nikon D90 camera with a Nikon AF 50 mm lens (Nikon Corporation, Tokyo, Japan). To measure cover, the camera was set to automatic exposure, Aperture-Priority and automatic autofocus mode and F10 lens. Following recommendations by Jonckheere et al. [-@Jonckheere2005], we used the automated Ridler algorithm to threshold the canopy photos. We then calculated canopy transmittance ($\tau_p$) from each canopy photo as the proportion of pixels with intensity in the blue band below that threshold. LAI from canopy photos (LAI~p~) was calculated with a simple Beer-Lambert law (Eq. 4).

@@

where $\tau_p$ canopy transmission estimated from canopy photos. We assumed that the extinction coefficient *k* = 0.5, following Macfarlane et al. [-@Macfarlane2007] recommendation for eucalypt forests. Without further calibration, the two estimates were reasonably well correlated, though on average LAI$_\tau$ was `r f(100*(with(flatcan_byring, mean(LAI.PAR.mean / LAI))-1))`% higher than $L_p$ (R^2^ = `r f(glance(lmphoto)$r.squared)`, RMSE = `r f(glance(lmphoto)$sigma)`) (Fig. S3). We also estimated the extinction coefficient that minimized the difference between the two estimates of LAI (Fig. S2), giving *k* = `r f(kopt)`. These comparisons are encouraging because the photos measure LAI near the zenith (+/- 15^o^), while $L_\tau$ is integrated over the sky hemisphere (Eq. 2), weighted towards the zenith.
